---
title: "God Spiked the Integers, and the Punch"
author: "Gina Nichols"
date: "8/18/2020"
output: html_document
---


```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE)
library(rethinking)
library(tidyverse)
library(cowplot)
library(emo)
theme_set(theme_bw())
```

# Easy

## E1

logodds = log(p/(1-p))

Lower chance than 50/50 is negative log-odds
If p = 0.35, logodds = `r round(log(0.35/(1-0.35)), 2)`

A 50% chance should mean a log-odds of 0 - the chances of the two are the same.
If p = 0.5, logodds = `r round(log(0.5/(1-0.5)), 2)`

Higher than 50/50 is positive logodds
If p = 0.75, logodds = `r round(log(0.75/(1-0.75)), 2)`

## E2

If logodds = 3.2, p = `r round(exp(3.2)/(1+exp(3.2)), 2)`

On some page he talks about this but I can't find it. 

## E3 `r emo::ji("help")`

I think it means that a unit increase in the predictor vastly increases the probability of the event. Maybe by  `r round(exp(1.7)/(1+exp(1.7)), 2)`%, but I'm not sure. 

## E4

Offsets make it possible to have things in different units. Say I measured weeds in a m-2 area, but my undergrad is a true American and she measured them in a ft-2 area. I can still use the data!

# Medium

## M1 `r emo::ji("help")`

Liklihood is related to entropy. In aggregated form of data, there are more ways to see each 'observation', so the entropy is higher, which I think changes what the likelihood looks like. 

## M2

One unit increase in the predictor results in an increase of almost 2 orders of magnitude in the response variable. 

## M3

It makes it so p is always between 0-1. **But it can never be 1, right? You can't divide by 0.**

## M4

The exponential of a value is always above 0, and counts are always above 0. 

## M5 `r emo::ji("help")`

Why would you use a logit link for the mean of a Poisson GLM? 
It means you expect the mean of a count variable to be between 0-1? No, I cannot think of a situation where you'd expect this. 

## M6 
*Constraints for which the binomial and Poisson have max entropy?*

Binomial - When 'a count outcome *y* for which each observation arises from *n* trials and with a constant expected value *np*.

Poisson - When the mean is equal to the variance. Seems more like a special case of the normal distribution, rather than the binomial distribution. But if you're in a binomial situation and n is very large, you have a Poisson

## M7 

*Create m11.4 using both quap and ulam*

```{r dat, include=F}

data("chimpanzees")
d <- chimpanzees %>% 
  mutate(trt = 1 + prosoc_left + 2*condition,
         treatment = trt)
```

```{r ulam, results="hide"}

dat_list <- 
  list(
    pulled_left = d$pulled_left,
    actor = d$actor,
    trt = as.integer(d$trt) )

m11.4 <- ulam(
  alist(
    pulled_left ~ dbinom(1, p),
    logit(p) <- a[actor] + b[trt],
    a[actor] ~ dnorm(0, 1.5),
    b[trt] ~ dnorm(0, 0.5) 
  ), data = dat_list, chains = 4, log_lik = T )
```


```{r quap}

m11.4quap <- quap(
  alist(
    pulled_left ~ dbinom(1, p),
    logit(p) <- a[actor] + b[trt],
    a[actor] ~ dnorm(0, 1.5),
    b[trt] ~ dnorm(0, 0.5) 
  ), data = d)

```



```{r}

rns <- rownames(precis(m11.4, depth = 2) )

precis(m11.4, depth = 2) %>% 
  as.data.frame() %>% 
  mutate(mod = "ulam") %>% 
  bind_rows(
    precis(m11.4quap, depth = 2) %>% 
      as.data.frame() %>% 
      mutate(mod = "quap")
  ) %>% 
  mutate(xparm = rep(rns, 2)) %>% 
  janitor::clean_names() %>% 
  ggplot(aes(mod, mean, color = mod)) + 
  geom_point() + 
  geom_linerange(aes(ymin = x94_5_percent, ymax = x5_5_percent)) +
  facet_grid(.~xparm) + 
  labs(title = "ulam and quap create almost identical estimates")
```

They are identical except for **old lefty**. Try using a terrible prior (normal(0, 10) for the actor intercept. 

```{r ulam2, results="hide"}

m11.4x <- ulam(
  alist(
    pulled_left ~ dbinom(1, p),
    logit(p) <- a[actor] + b[trt],
    a[actor] ~ dnorm(0, 10),
    b[trt] ~ dnorm(0, 0.5) 
  ), data = dat_list, chains = 4, log_lik = T )
```
```{r}

m11.4quapx <- quap(
  alist(
    pulled_left ~ dbinom(1, p),
    logit(p) <- a[actor] + b[trt],
    a[actor] ~ dnorm(0, 10),
    b[trt] ~ dnorm(0, 0.5) 
  ), data = d)

rns <- rownames(precis(m11.4x, depth = 2) )

precis(m11.4x, depth = 2) %>% 
  as.data.frame() %>% 
  mutate(mod = "ulam") %>% 
  bind_rows(
    precis(m11.4quapx, depth = 2) %>% 
      as.data.frame() %>% 
      mutate(mod = "quap")
  ) %>% 
  mutate(xparm = rep(rns, 2)) %>% 
  janitor::clean_names() %>% 
  ggplot(aes(mod, mean, color = mod)) + 
  geom_point() + 
  geom_linerange(aes(ymin = x94_5_percent, ymax = x5_5_percent)) +
  facet_grid(.~xparm) + 
  labs(title = "ulam had a terrible time with that prior and old lefty")
```

Why does MCMC have a harder time? It goes wandering, I guess. 

# M8

Use the Kline data but drop Hawaii. 

```{r, results="hide"}
data(Kline)
d2 <- Kline %>% 
  mutate(P = scale(log(population)),
         contact_id = ifelse(contact == "high", 2, 1))

dat2 <- list(
  T = d2$total_tools,
  P = d2$P,
  cid = d2$contact_id
)

m11.10 <- ulam(
  alist(
    T ~ dpois( lambda ),
    log(lambda) <- a[cid] + b[cid]*P,
    a[cid] ~ dnorm( 3, 0.5 ),
    b[cid] ~ dnorm( 0, 0.2 )
  ), data = dat2, chains = 4, log_lik = TRUE )
```


```{r}
d2out <- Kline %>% 
  mutate(P = scale(log(population)),
         contact_id = ifelse(contact == "high", 2, 1)) %>% 
  filter(culture != "Hawaii")

dat2out <- list(
  T = d2out$total_tools,
  P = d2out$P,
  cid = d2out$contact_id
)

m11.10out <- ulam(
  alist(
    T ~ dpois( lambda ),
    log(lambda) <- a[cid] + b[cid]*P,
    a[cid] ~ dnorm( 3, 0.5 ),
    b[cid] ~ dnorm( 0, 0.2 )
  ), data = dat2out, chains = 4, log_lik = TRUE )
```

```{r}

rns2 <- rownames(precis(m11.10, depth = 2) )
rns3 <- rownames(precis(m11.10out, depth = 2) )

precis(m11.10, depth = 2) %>% 
  as.data.frame() %>% 
  mutate(mod = "HI") %>% 
  bind_rows(
    precis(m11.10out, depth = 2) %>% 
      as.data.frame() %>% 
      mutate(mod = "noHI")
  ) %>% 
  mutate(xparm = c(rns2, rns3)) %>% 
  janitor::clean_names() %>% 
  ggplot(aes(mod, mean, color = mod)) + 
  geom_point() + 
  geom_linerange(aes(ymin = x94_5_percent, ymax = x5_5_percent)) +
  facet_grid(.~xparm) + 
  labs(title = "1 is low contact, 2 is high. Makes sense 1 is affected more by removal.")
```

# Hard

## H4

```{r}
data("NWOGrants")
gr <-
  NWOGrants %>% 
  mutate(gid = ifelse(gender == "m", 1, 2),
         deptid = as.numeric(as.factor(discipline)))

gr %>% 
  ggplot(aes(gid, awards)) + 
  geom_line(color = "red") + 
  geom_line(aes(y = applications), color = "black") + 
  scale_x_continuous(breaks = c(1, 2), labels = c("M", "F")) +
  facet_grid(.~discipline) + 
  labs(title = "Females apply for less grants in some fields")
```

```{r}

mh4 <- quap(
  alist(
    awards ~ dbinom( applications, p ),
    logit(p) <- a[gid],
    a[gid] ~ dnorm(0, 1.5)
  ), data = gr )

precis(mh4, depth = 2)
```

Look at differences on scale that makes sense
```{r}
post <- extract.samples(mh4)
diff_a <- post$a[,1] - post$a[,2]
diff_p <- inv_logit(post$a[,1]) - inv_logit(post$a[,2])
precis(list (diff_a = diff_a, diff_p= diff_p))
```

Men are 1-5% more likely to be funded. 


But what if we ask if it's different within fields?

```{r}

mh4b <- quap(
  alist(
    awards ~ dbinom( applications, p ),
    logit(p) <- a[gid] + delta[deptid],
    a[gid] ~ dnorm(0, 1.5),
    delta[deptid] ~ dnorm(0, 1.5)
  ), data = gr )

precis(mh4b, depth = 2)
post <- extract.samples(mh4b)
diff_a <- post$a[,1] - post$a[,2]
diff_p <- inv_logit(post$a[,1]) - inv_logit(post$a[,2])
precis(list (diff_a = diff_a, diff_p= diff_p))
```

Look at things sorted by overall applications

```{r}
gr %>% 
  group_by(discipline) %>% 
  mutate(totaps = sum(applications),
         pctaw = awards/totaps) %>% 
  arrange(totaps) %>% 
  ggplot(aes(gid, pctaw)) + 
  geom_point(aes(size = totaps, color = gender)) +
  scale_x_continuous(breaks = c(1, 2), labels = c("M", "F"), expand = c(1, 1)) +
  facet_grid(.~discipline) + 
  labs(title = "Females apply for less grants in some fields")

gr %>% 
  group_by(discipline) %>% 
  mutate(totaps = sum(applications),
         pctaw = awards/totaps) %>% 
  arrange(totaps) %>% 
  ggplot(aes(reorder(discipline, totaps), pctaw)) + 
  geom_point(aes(size = totaps, color = gender)) +
  #scale_x_continuous(breaks = c(1, 2), labels = c("M", "F"), expand = c(1, 1)) +
#  facet_grid(.~discipline) + 
  labs(title = "Females apply for less grants in some fields") + 
  theme(axis.text.x = element_text(angle = 45))


```

