---
title: "Chapter 11 Stat Rethink"
author: "Anabelle Laurent"
date: "8/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rethinking)
library(emo)
library(dplyr)
library(ggplot2)
```


### M.7 `r emo::ji("monkey")`

```{r}
data(chimpanzees)
d <- chimpanzees
d$block_id <- d$block
d$treatment <- 1L + d$prosoc_left + 2L*d$condition
dat_list <- list(
    pulled_left = d$pulled_left,
    actor = d$actor,
    block_id = d$block,
    treatment = as.integer(d$treatment) )

m11.4ulam <- ulam(
    alist(
        pulled_left ~ dbinom( 1 , p ) ,
        logit(p) <- a[actor] + b[treatment] ,
        a[actor] ~ dnorm( 0 , 1.5 ),
        b[treatment] ~ dnorm( 0 , 0.5 )
    ) , data=dat_list , chains=4 , log_lik=TRUE )


m11.4quap<-quap(
    alist(
        pulled_left ~ dbinom( 1 , p ) ,
        logit(p) <- a[actor] + b[treatment] ,
        a[actor] ~ dnorm( 0 , 1.5 ),
        b[treatment] ~ dnorm( 0 , 0.5 )
    ) , data=dat_list )
precis( m11.4ulam , depth=2 )
precis( m11.4quap , depth=2 )
```
The individual intercepts are similar between quap and ulam.  

```{r}
m11.4ulam10 <- ulam(
    alist(
        pulled_left ~ dbinom( 1 , p ) ,
        logit(p) <- a[actor] + b[treatment] ,
        a[actor] ~ dnorm( 0 , 10 ),
        b[treatment] ~ dnorm( 0 , 0.5 )
    ) , data=dat_list , chains=4 , log_lik=TRUE )


m11.4quap10<-quap(
    alist(
        pulled_left ~ dbinom( 1 , p ) ,
        logit(p) <- a[actor] + b[treatment] ,
        a[actor] ~ dnorm( 0 , 10 ),
        b[treatment] ~ dnorm( 0 , 0.5 )
    ) , data=dat_list )
precis( m11.4ulam10 , depth=2 )
precis( m11.4quap10 , depth=2 )
```

By relaxing the prior we get a very large credible interval for the chimpazee 2!

### H.1

```{r}
m11.4ulam <- ulam(
    alist(
        pulled_left ~ dbinom( 1 , p ) ,
        logit(p) <- a[actor] + b[treatment] ,
        a[actor] ~ dnorm( 0 , 1.5 ),
        b[treatment] ~ dnorm( 0 , 0.5 )
    ) , data=dat_list , chains=4 , log_lik=TRUE )

m11.3ulam <- ulam(
    alist(
        pulled_left ~ dbinom( 1 , p ) ,
        logit(p) <- a + b[treatment] ,
        a ~ dnorm( 0 , 1.5 ),
        b[treatment] ~ dnorm( 0 , 0.5 )
    ) , data=dat_list ,chains=4 , log_lik=TRUE,iter=2000)

compare( m11.4ulam , m11.3ulam , func=WAIC )
```
ESS was too low, I increased the number of iterations.  
The model with the individual intercept has the lowest WAIC.  

### H2.a `r emo::ji("eagle")`

```{r}
library(MASS)
data(eagles)
pir <- eagles
head(eagles)
```
First we need to create dummy variables with 1 for large, 0 for small and 1 for adult, 0 for immature.  
We need to transform the variables P (size of pirate `r emo::ji("pirate")`) ,A (age of the pirate) and V (size of the victim).  
y = number of success  
n = total number of trials  
Then the probability is Y/n  

```{r}
library(fastDummies)
pir2<-dummy_cols(pir, select_columns = c("P", "A","V"))
head(pir2)
pir2$success<-pir2$y/pir2$n
```

```{r}
mod_pirate <- quap(alist(
    y ~ dbinom(n, p),
    logit(p) <- a + bp*P_L + ba*A_A + bv*V_L,
    a ~ dnorm(0, 10),
    bp ~ dnorm(0, 5),
    bv ~ dnorm(0, 5),
    ba ~ dnorm(0, 5)
  ), data=pir2)
precis(mod_pirate, depth=2 )

mod_pirate_ulam <- ulam(alist(
    y ~ dbinom(n, p),
    logit(p) <- a + bp*P_L + ba*A_A + bv*V_L,
    a ~ dnorm(0, 10),
    bp ~ dnorm(0, 5),
    bv ~ dnorm(0, 5),
    ba ~ dnorm(0, 5)
  ), data=pir2,chains=4 , log_lik=TRUE )
precis(mod_pirate_ulam, depth=2 )
plot(precis(mod_pirate_ulam))
```

### H2b

If the victim is large, it negatively impact the number of successful attemps (bv parameter).  
parameters bp and ba positively impact the number of successful attemps.  


```{r}
post <- extract.samples(mod_pirate_ulam)
str(post)
success <- inv_logit( post$a )
plot( precis( as.data.frame(success) ) , xlim=c(0,1) )
```

Let's look at the observed data first
```{r}
pir2<-pir2 %>% mutate(category=paste0(P,A,V))
pir2 %>% ggplot(aes(y=success,x=category,col="sim")) +geom_point() + theme_bw() 
  
```

If an eagle is small + immature, he can keep trying `r emo::ji("laugh")`

```{r}
p_post <- link( mod_pirate_ulam )
p_mu <- apply( p_post , 2 , mean )
p_ci <- apply( p_post , 2 , PI )

pir2 %>% ggplot(aes(y=success,x=category,col="sim")) +geom_point() + theme_bw() +
  geom_point(y=p_mu, aes(col="pred")) 
    scale_color_manual("Legend",
                     values=c("red","blue"), 
                     breaks = c("sim", "pred"),
                     labels=c("observed success","predicted success"))

```

### H2.c

```{r}
mod_inter <- ulam(alist(
    y ~ dbinom(n, p),
    logit(p) <- a + bp*P_L + ba*A_A + bv*V_L + bpa*P_L*A_A,
    a ~ dnorm(0, 10),
    bp ~ dnorm(0, 5),
    bv ~ dnorm(0, 5),
    ba ~ dnorm(0, 5),
    bpa ~ dnorm(0, 5)
  ), data=pir2,chains=4 , log_lik=TRUE )
precis(mod_inter, depth=2 )
plot(precis(mod_inter))


compare( mod_pirate_ulam , mod_inter , func=WAIC )
```

