---
title: "Stat_rethink Chapter 9"
author: "Mariana CHiozza"
date: "8/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# God Spiked the Integers


## 11.1 Binomial Regression

### 11.1.1 Logistic regression.

Prosocial chimpanzees example.

HOW MUCH DIFFERENCE DOES THE TREATMENT MAKE IN THE PROBABILITY OF PULLING A LEVER?

```{r}
library(rethinking)
data(chimpanzees)
d <- chimpanzees

d$treatment <- 1 + d$prosoc_left + 2*d$condition
```

Regularized priors and unknown posterior:


```{r}

# trimmed data list
dat_list <- list(
    pulled_left = d$pulled_left,
    actor = d$actor,
    treatment = as.integer(d$treatment) )


m11.4 <- ulam(
    alist(
        pulled_left ~ dbinom( 1 , p ) ,
        logit(p) <- a[actor] + b[treatment] ,
        a[actor] ~ dnorm( 0 , 1.5 ),
        b[treatment] ~ dnorm( 0 , 0.5 )
    ) , data=dat_list , chains=4 , log_lik=TRUE )
precis( m11.4 , depth=2 )

```


### 11.1.2 Relative shark and absolute deer.

PROPORTIONAL ODDS OF SWITCHING FROM ONE TREATMENT TO ANOTHER BY EXPONENTIATING THE PARAMETER OF INTEREST.


### 11.1.3 Aggregated binomial.


```{r}

data(chimpanzees)
d <- chimpanzees
d$treatment <- 1 + d$prosoc_left + 2*d$condition
d$side <- d$prosoc_left + 1 # right 1, left 2
d$cond <- d$condition + 1 # no partner 1, partner 2
d_aggregated <- aggregate(
    d$pulled_left ,
    list( treatment=d$treatment , actor=d$actor ,
          side=d$side , cond=d$cond ) ,
    sum )
colnames(d_aggregated)[5] <- "left_pulls"


dat <- with( d_aggregated , list(
    left_pulls = left_pulls,
    treatment = treatment,
    actor = actor,
    side = side,
    cond = cond ) )

m11.6 <- ulam(
    alist(
        left_pulls ~ dbinom( 18 , p ) ,
        logit(p) <- a[actor] + b[treatment] ,
        a[actor] ~ dnorm( 0 , 1.5 ) ,
        b[treatment] ~ dnorm( 0 , 0.5 )
    ) , data=dat , chains=4 , log_lik=TRUE )

precis(m11.6, depth = 2)

```

TO CALCULATE WAIC OR PSIS, USE A LOGISTIC REFRESSION DATA FORMAT, NOT AN AGGREGATED FORMAT.


GRADUATE SCHOOL ADMISIONS EXAMPLE, TO SHOW HOW ADDING A FACTOR TO THE MODEL (DEPARTMENTS) CHANGE THE INFERENCE ABOUT HOW GENDER AFFECT THE ADMISSIONS.

Q1: WHAT ARE THE AVERAGE PROBABILITIES OF ADMISSION FOR WOMEN AND MEN ACROSS ALL DEPARTMENTS?

Q2: WHAT IS THE AVERAGE DIFFERENCE IN PROBABILITY OF ADMISSION BETWEEN WOMEN AND MEN WITHIN DEPARTMENTS?

DEPARTMENT IS A CONFOUND. CONTROLLING THE DEPARTMENT REVEALS A MORE PLAUSIBLE DIRECT CAUSAL INFLUENCE OF GENDER.

## Poisson Distribution

When N is unknown or uncoutably large.

Only 1 parameter, which is the expected value and the variance.

### Negative binomial


