---
title: "Statistical Rethinking Chapter 3 Exercises"
author: "Anabelle Laurent"
date: "6/6/2020"
output:  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rethinking)
devtools::install_github("hadley/emo")
library(emo)
```

## Practice

```{r}
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prior <- rep( 1 , 1000 )
likelihood <- dbinom( 6 , size=9 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)
set.seed(100)
samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )
plot(samples)
dens(samples)
```


## E.1
```{r}
# 3E1 add up posterior probability where p<0.2
sum(samples<0.2)/1e4
```

## E.2
```{r}
sum( samples > 0.8) / 1e4
```

## E.3
```{r}
sum( samples > 0.2 & samples < 0.8 ) / 1e4
```

## E.4
```{r}
quantile(samples,0.2)
```

## E.5
```{r}
quantile(samples,0.8)
```

## E.6
```{r}
HPDI(samples,prob=0.66)
```

## E.7
```{r}
PI(samples, prob=0.66)
```

## M.1
```{r}
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prior <- rep( 1 , 1000 )
likelihood <- dbinom( 8 , size=15 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)
plot(p_grid,posterior)
```

## M.2
```{r}
set.seed(100)
samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )
HPDI(samples,prob=0.90)
```

## M.3
We use samples from the posterior distribution used in 3.M2  
The object w can be manipulated just like samples from the distribution 

```{r}
w <- rbinom( 1e5 , size=15 , prob=samples )
simplehist( w , xlab="water count" )

```

What is the proba of observing 8 water in 15 tosses?
```{r}
sum( w==8) / length(w)
```

## M.4
```{r}
w <- rbinom( 1e5 , size=6 , prob=samples )
simplehist( w , xlab="water count" )
sum( w==6) / length(w)
```

## M.5 
```{r}
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prior <- c(rep( 0 , 500 ),rep( 1 , 500 ))
likelihood <- dbinom( 8 , size=15 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)
plot(p_grid,posterior,type="l")
```

## M.6

## Hard problems `r emo::ji("deadpan")`

```{r}
birth1 <- c(1,0,0,0,1,1,0,1,0,1,0,0,1,1,0,1,1,0,0,0,1,0,0,0,1,0,
            0,0,0,1,1,1,0,1,0,1,1,1,0,1,0,1,1,0,1,0,0,1,1,0,1,0,0,0,0,0,0,0,
            1,1,0,1,0,0,1,0,0,0,1,0,0,1,1,1,1,0,1,0,1,1,1,1,1,0,0,1,0,1,1,0,
            1,0,1,1,1,0,1,1,1,1)
birth2 <- c(0,1,0,1,0,1,1,1,0,0,1,1,1,1,1,0,0,1,1,1,0,0,1,1,1,0,
            1,1,1,0,1,1,1,0,1,0,0,1,1,1,1,0,0,1,0,1,1,1,1,1,1,1,1,1,1,1,1,1,
            1,1,1,0,1,1,0,1,1,0,1,1,1,0,0,0,0,0,0,1,0,0,0,1,1,0,0,1,0,0,1,1,
            0,0,0,1,1,1,0,0,0,0)

data(homeworkch3)
# total number of boys across all of these births
sum(birth1)+sum(birth2)
```

1 means `r emo::ji("baby")` boy  
2 means `r emo::ji("baby")` girl  

Total number of boys across all of these births  
```{r}
sum(birth1)+sum(birth2)
```
 
## H.1
```{r}
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prior <- rep( 1 , 1000 )
likelihood <- dbinom( 111 , size=200 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)
plot(p_grid,posterior)
```

## H.2
```{r}
set.seed(100)
samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )
HPDI(samples,prob=c(0.5,0.89,0.97))
```

## H.3
```{r}
w <- rbinom( 1e4 , size=200 , prob=samples )
dens(w) 
abline(v=111)
```

## H.4
```{r}
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prior <- rep( 1 , 1000 )
likelihood <- dbinom( sum(birth1) , size=100 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)
plot(p_grid,posterior)
set.seed(100)
samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )
w <- rbinom( 1e4 , size=100 , prob=samples )
dens(w) 
abline(v=sum(birth1))
```

## H.5 (not sure about this one)

First count the number of first borns who were girls
```{r}
girl_birth1<-100-sum(birth1)
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prior <- rep( 1 , 1000 )
likelihood <- dbinom( girl_birth1 , size=100 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)
plot(p_grid,posterior)
set.seed(100)
samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )
w <- rbinom( 1e4 , size=100 , prob=samples )
dens(w) 
abline(v=sum(birth1))
```


