---
title: "Monsters and Mixtures"
author: "Gina Nichols"
date: "8/25/2020"
output: html_document
---


```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE)
library(rethinking)
library(tidyverse)
library(patchwork)
library(emo)
theme_set(theme_bw())
```

# Easy

## E1 Ordered categorical vs categorical
Categorical variables like gender (m/f/other) or race have no order. Ordered categorical variables have a natural sequence to them, but the distance between entries might not be equal. Depth of sleep might be an example. Light sleep, medium, and deep. 

## E2 Link function of ordered logistic regression
Ordered logistic regression employs a the log-cumulative odds. 

## E3 What kind of inferential error do you make if you ignore zero-inflation? `r emo::ji("help")`
You will underestimate the mean rate?

## E4 Example of over- and under-dispersed counts `r emo::ji("help")`
Heterogeneity in observations by some unknown (or known?) factor can cause over-dispersed counts. For example if you count the number of raccoons in your yard and don't account for your neighbor having a bbq and throwing out lots of chicken bones, you might get over-dispersion (more high values than you would've expected). **Could zero-inflation cause under-dispersed counts?**

# Medium

## M1 and M2 Compute log-cum-odds of prof productivity

```{r}
uni_prod <- tibble(
  rating = c(1, 2, 3, 4),
  tot = c(12, 36, 7, 41)
) %>% 
  mutate(
    prop = tot/sum(tot)
  )

m1_hist <- 
  uni_prod %>% 
  ggplot(aes(rating, tot)) +
  geom_col() + 
  labs(title = "Super productive employees")

m1_cumprop <- 
  uni_prod %>% 
  mutate(
    cum_prop = cumsum(prop)
  ) %>% 
  ggplot(aes(rating, cum_prop)) + 
  geom_point() + 
  geom_line() + 
  labs(title = "Cumulative Proportions")

m1_logcumodds <- 
  uni_prod %>% 
  mutate(
    cum_prop = cumsum(prop),
    cum_odds = cum_prop/(1 - cum_prop),
    log_cum_odds = log(cum_odds)
  ) %>% 
  ggplot(aes(rating, log_cum_odds)) + 
  geom_point() + 
  geom_line() + 
  labs(title = "Log-Cumulative-Odds")

m1_hist + m1_cumprop + m1_logcumodds

```

## M3 Can you construct a zero-inflated binomial distribution? `r emo::ji("help")`

No. 