---
title: "Chap 12 Stat Rethink"
author: "Anabelle Laurent"
date: "8/21/2020"
output: html_document
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE)
library(rethinking)
library(tidyverse)
library(cowplot)
library(emo)
theme_set(theme_bw())
```

 results="hide"

When the observed variance exceed *Np(1-p)* with *Np* beign the expected value of a binomial, then we have overdispersion. It implies that omitted variable(s) is producing additional dispersion to the observed counts.  
<p style="color:orange">Question 1: is the omitted variable the only reason to produce overdipersion? What about biological variation, error of measurement?</p>

**Beta-binomial** estimates the distribution of probabilities of success instead of a single probability of success.  

**Cumulative probability** is the probability of a value or *any smaller value*.  



```{r}
pbar <- 0.5
theta <- 2
curve( dbeta2(x,pbar,theta) , from=0 , to=1 ,
    xlab="Different possible probability values" , ylab="Density" )
mtext("theta=2")
```
When theta (shape parameter that describes how spread out is the distribution) =2, every proba from 0 to 1 is equally likely.  

```{r}
pbar <- 0.5
theta <- 5
curve( dbeta2(x,pbar,theta) , from=0 , to=1 ,
    xlab="Different possible probability values" , ylab="Density" )
mtext("theta=5")
```

When theta >2, the distribution of proba grows more concentrated.  

```{r}
pbar <- 0.5
theta <- 1
curve( dbeta2(x,pbar,theta) , from=0 , to=1 ,
    xlab="Different possible probability values" , ylab="Density" )
mtext("theta=1")
```

When theta<1, extreme proba near 0 an 1 are more likely than the mean probability.  


<p style="color:orange">Question 2: Why binomial does not return a distribution? With the chain we can get a distribution of the outcome</p>  


**Negative-binomial or gamma-Poisson** assumes that each Poisson count has its own rate. As Gamma-poisson expected more variation around the mean, a high influential data in Poisson becomes less influential with Gamma-Poisson.  

##E4
Overdispersion: you go to a shelter and you ask dogs to seat (n=10) (if the dog seats=success). As dogs in a shelter varies in age, sex, breed, history you can expect overdispersion.  
Underdispersion: you go to a dog breeder already educating his dogs for adoption, you can expect a success most of the time.  

## M1 & M2 
 

```{r}
rate<-c(rep(1,12),rep(2,36),rep(3,7),rep(4,41))

# discrete proportion of each response value
pr_k <- table( rate ) / 96

# cumsum converts to cumulative proportions
cum_pr_k <- cumsum( pr_k )

logit <- function(x) log(x / (1 - x)) # convenience function
logit(cum_pr_k)
plot(1:4,logit(cum_pr_k),type="b",xlab="response",ylab="log-cumulative-odds")

dplot<-data.frame(rate=c(1:4),cum_pr_k)

dplot %>% ggplot() + 
  geom_line(aes(x=c(1,2,3,4),y=cum_pr_k)) +
  geom_point(aes(x=c(1,2,3,4),y=cum_pr_k),size=2) +
  xlab("response") +
  ylab("cumulative proportion")
```

## H1 `r emo::ji("tornado")`

```{r}
data(Hurricanes)
d <- Hurricanes

d %>% 
  ggplot(aes(x = category)) +
  geom_histogram(binwidth = 1/4, size = 0) +
  xlab("Hurricane category") +
  ylab("Frequency")

d2<- d %>% group_by(category,female) %>% summarise (n=sum(deaths))
ggplot(d2,aes(category,n,fill=as.factor(female)))+
  geom_bar(stat="identity",position='dodge') +
  ylab("number of deaths")

d3<- d %>% group_by(femininity) %>% summarise (n=sum(deaths)) 
ggplot(d3,aes(x=femininity,y=n))+ geom_point() +
   ylab("number of deaths") +
  xlab("femininity scale")
```


```{r, results="hide"}
dat <- list(
    D = d$deaths ,
    F = d$femininity)

# intercept only
m1<- ulam(
    alist(
        D ~ dpois( lambda ),
        log(lambda) <- a,
        a ~ dnorm( 0 , 10)
    ), data=dat , chains=4 , log_lik=TRUE )


#  model with predictor
m2 <- ulam(
    alist(
        D ~ dpois( lambda ),
        log(lambda) <- a + b*F,
        a~ dnorm( 0 , 10 ),
        b~ dnorm( 0 , 5 )
    ), data=dat , chains=4 , log_lik=TRUE )
precis(m2, depth = 2)
```
```{r}
precis(m1, depth = 2)
precis(m2, depth = 2)
compare( m1 , m2 , func=PSIS )
```


## H2

```{r results="hide"}

m3<- ulam(
    alist(
        D ~ dgampois( lambda , phi ),
        log(lambda) <- a + b*F,
        a ~ dnorm(0,10),
        b ~ dnorm(0,5),
        phi ~ dexp(1)
    ), data=dat , chains=4 , log_lik=TRUE )
```

```{r}
precis(m3,depth = 2)
compare(m2,m3,func=PSIS)
```

