---
title: "Ulysses' Compass"
author: "Gina Nichols"
date: "7/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(rethinking)
```

# Chapter 7 Practice


## 7E1.

*Three motivating criteria defining information entropy*

1. We want our metric to be continuous. 
2. It should have a higher value when there are more possible events. 
3. We want the calculation of it to be symmetrical. ie It shouldn't matter how we break down the event combinations, they should always add up to the same value. 

## 7E2. 
```{r}
p2 <- c(0.7, 0.3)
round(-sum(log(p2)*p2), 2)
```

Entropy is `r round(-sum(log(p2)*p2), 2)`

## 7E3. 
```{r}
p3 <- c(0.2, 0.25, 0.25, 0.3)
round(-sum(log(p3)*p3), 2)
```

Entropy is `r round(-sum(log(p3)*p3), 2)`

## 7E4. 
```{r}
p4 <- c(1/3, 1/3, 1/3)
round(-sum(log(p4)*p4), 2)
```

Entropy is `r round(-sum(log(p4)*p4), 2)`


Curiosity
```{r}
x <- seq(0, 1, by =0.1)
y <- 1-x

tibble(x = x,
       y = y,
       entx = x*log(x),
       enty = y*log(y)) %>% 
  mutate(entx = ifelse(is.nan(entx), 0, entx),
         enty = ifelse(is.nan(enty), 0, enty),
         ent = -(entx + enty)) %>% 
  ggplot(aes(x, y)) +
  geom_point(aes(color = ent), size = 5) + 
  scale_color_viridis_c()

```

## 7M1. 

$AIC = D_{train} + 2p = -2lppd + 2p$
$WAIC = -2lppd + 2\sum var_{\theta}\log p(y_i|\theta)$

It seems like the penalty term is the only difference. Instad of just the number of parameters $p$ in $AIC$, we have an "effective" number of parameters. 

I don't quite understand what the *lppd* is. 
Log-posterior-predictive-density.
The log of the average probability for each observation...?


