---
title: "Stat_rethink Chapter 15"
author: "Mariana CHiozza"
date: "9/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Missing Data and Other Opportunities

## Practice Problems

### 15E1.


Ti ~ Poisson (mu_i) 

logmu_i = alpha + beta*logP_True_i

logP_Obs_i ~ Normal (logP_True_i, logP_SE_i)

logP_True_i ~ Normal (0, 1)

alpha ~ Normal(0, 1.5)

beta ~ Normal(0, 1)


### 15E2.

Ti ~ Poisson (mu_i) 

logmu_i = alpha + beta*logP_i

logP_i ~ Normal (nu, sigma_logP_i)

nu ~ Normal (0, 0.5)

sigma_logP_i ~ Exponential (1)

alpha ~ Normal(0, 1.5)

beta ~ Normal(0, 1)
    


### 15M1.

The model assumes that each parameter for the imputation of each logP_i has a standarized Gaussian uncertainty. 



### 15M2.

Can't get the beta distribution for B......


```{r}
## R code 15.16
library(rethinking)
data(milk)
d <- milk
d$neocortex.prop <- d$neocortex.perc / 100
d$logmass <- log(d$mass)
dat_list <- list(
    K = standardize( d$kcal.per.g ),
    B = standardize( d$neocortex.prop ),
    M = standardize( d$logmass ) )

## R code 15.17
m15.5 <- ulam(
    alist(
        K ~ dnorm( mu , sigma ),
        mu <- a + bB*B + bM*M,
        B ~ dnorm( nu , sigma_B ),
        c(a,nu) ~ dnorm( 0 , 0.5 ),
        c(bB,bM) ~ dnorm( 0, 0.5 ),
        sigma_B ~ dexp( 1 ),
        sigma ~ dexp( 1 )
    ) , data=dat_list , chains=4 , cores=4 )

## R code 15.18
precis( m15.5 , depth=2 )


```


### 15M3.

```{r}
## R code 15.2
library(rethinking)
data(WaffleDivorce)
d <- WaffleDivorce

# points
plot( d$Divorce ~ d$MedianAgeMarriage , ylim=c(4,15) ,
    xlab="Median age marriage" , ylab="Divorce rate" )

# standard errors
for ( i in 1:nrow(d) ) {
    ci <- d$Divorce[i] + c(-1,1)*d$Divorce.SE[i]
    x <- d$MedianAgeMarriage[i]
    lines( c(x,x) , ci )
}

## R code 15.5
dlist <- list(
    D_obs = standardize( d$Divorce ),
    D_sd = d$Divorce.SE / sd( d$Divorce ),
    M_obs = standardize( d$Marriage ),
    M_sd = d$Marriage.SE / sd( d$Marriage ),
    A = standardize( d$MedianAgeMarriage ),
    N = nrow(d)
)

m15.2 <- ulam(
    alist(
        D_obs ~ dnorm( D_true , D_sd ),
        vector[N]:D_true ~ dnorm( mu , sigma ),
        mu <- a + bA*A + bM*M_true[i],
        M_obs ~ dnorm( M_true , M_sd ),
        vector[N]:M_true ~ dnorm( 0 , 1 ),
        a ~ dnorm(0,0.2),
        bA ~ dnorm(0,0.5),
        bM ~ dnorm(0,0.5),
        sigma ~ dexp( 1 )
    ) , data=dlist , chains=4 , cores=4 )

## R code 15.6
post <- extract.samples( m15.2 )
D_true <- apply( post$D_true , 2 , mean )
M_true <- apply( post$M_true , 2 , mean )
plot( dlist$M_obs , dlist$D_obs , pch=16 , col=rangi2 ,
    xlab="marriage rate (std)" , ylab="divorce rate (std)" )
points( M_true , D_true )
for ( i in 1:nrow(d) )
    lines( c( dlist$M_obs[i] , M_true[i] ) , c( dlist$D_obs[i] , D_true[i] ) )



```

```{r}
## R code 15.5 modif
dlist.2 <- list(
    D_obs = standardize( d$Divorce ),
    D_sd = 2*(d$Divorce.SE) / sd( d$Divorce ),
    M_obs = standardize( d$Marriage ),
    M_sd = 2*(d$Marriage.SE) / sd( d$Marriage ),
    A = standardize( d$MedianAgeMarriage ),
    N = nrow(d)
)

m15.2.2 <- ulam(
    alist(
        D_obs ~ dnorm( D_true , D_sd ),
        vector[N]:D_true ~ dnorm( mu , sigma ),
        mu <- a + bA*A + bM*M_true[i],
        M_obs ~ dnorm( M_true , M_sd ),
        vector[N]:M_true ~ dnorm( 0 , 1 ),
        a ~ dnorm(0,0.2),
        bA ~ dnorm(0,0.5),
        bM ~ dnorm(0,0.5),
        sigma ~ dexp( 1 )
    ) , data=dlist.2 , chains=4 , cores=4 )

## R code 15.6
post <- extract.samples( m15.2.2 )
D_true <- apply( post$D_true , 2 , mean )
M_true <- apply( post$M_true , 2 , mean )
plot( dlist.2$M_obs , dlist.2$D_obs , pch=16 , col=rangi2 ,
    xlab="marriage rate (std)" , ylab="divorce rate (std)" )
points( M_true , D_true )
for ( i in 1:nrow(d) )
    lines( c( dlist.2$M_obs[i] , M_true[i] ) , c( dlist.2$D_obs[i] , D_true[i] ) )

```



