---
title: "Geocentric Models, Practice"
author: "Gina Nichols"
date: "6/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rethinking)
#library(brms)
```

**I'd like to discuss M6 and M7**

### E1 Which of these is the likelihood?
$$
\begin{array}{l}
y_i \sim Normal(\mu, \sigma) \\
\mu \sim Normal(0, 10) \\
\sigma \sim Exponential(1) \\
\end{array}
$$

$y_i \sim Normal(\mu, \sigma)$ is the likelihood.

### E2 How many parm in prior?
There are 2 parameters in the posterior distribution, $\mu$ and $\sigma$.

### E3 Write that model using Bayes' theorem. 

Ugh. Not that good in Latex.

### E4 Which of these is the linear model?
$$
\begin{array}{l}
h_i \sim Normal(\mu_i, \sigma) \\
\mu_i = \alpha + \beta (x_i - \bar{x}) \\
\alpha \sim Normal(0, 10) \\
\beta \sim Normal(0, 1) \\
\sigma \sim Exponential(2) \\
\end{array}
$$

$\mu_i = \alpha + \beta (x_i - \bar{x})$ is our linear model. Although I think there is a typo in the book in the likelihood. 

### E5 How many parameters are in that posterior distribution?
Three - $\alpha$, $\beta$, and $\sigma$. 

### M1 Simulate observed y values from the prior for the model below:

$$
\begin{array}{l}
y_i \sim Normal(\mu, \sigma) \\
\mu \sim Normal(0, 10) \\
\sigma \sim Exponential(1) \\
\end{array}
$$

```{r}
sample_mu <- rnorm (1e4, 0, 10)
sample_sig <- rexp(1e4, 1)
prior_h <- rnorm(1e4, sample_mu, sample_sig)
dens(prior_h)
```

### M2 Translate this into a quap formula.

```{r}
flist <- 
  alist(
    height ~ dnorm(mu, sigma),
    mu ~ dnorm(0, 10),
    sigma ~ dexp(1)
  )
```


### M3 
This is the same as **E4**

### M4 Mathematical model for student heights over 3 years
You measure student heights over 3 years. I want a linear regression predicting height using year as a predictor. 
Write a mathematical model for this. Defend it!

$$
\begin{array}{l}
h_i \sim Normal(\mu_i, \sigma) \\
\mu_i = \alpha + \beta (year_i - \bar{year}) \\
\alpha \sim Normal(0, 10) \\
\beta \sim LogNormal(0, 1) \\
\sigma \sim Exponential(1) \\
\end{array}
$$

### M5 Think about the priors
No because I already forced the $\beta$ to be positive by using that LogNormal thing. 

### M6 Update priors based on new info
I'm not sure how I'd add this information to the *exponential* distribution. In my ignorance I suppose I'd update it to:

$\sigma \sim Uniform(0, 8)$

**I'd like help on this one**

### M7 Height as function of weight, HELP 
Fit m4.3 again. 

```{r}
data(Howell1)
d2 <-  
  Howell1 %>% 
  filter(age >= 18)

xbar <- mean(d2$height)

hmod_cent <- 
  quap(
    alist(
      height ~ dnorm(mu, sigma),
      mu <- a + b*(weight - xbar),
      a ~ dnorm(178, 20),
      b ~ dlnorm(0, 1),
      sigma ~ dunif(0, 50)
    ),
    data = d2)

precis(hmod_cent)
```

Hmmm. I'm not getting the same answer they got. What is wrong?


Pretend it's fine, I guess. 
```{r}

post_hmod_cent <- extract.samples(hmod_cent) #--get our posterior samples of a, b, and sigma

p_hmod_cent <- 
  d2 %>% 
  mutate(a = mean(post_hmod_cent$a),
         b = mean(post_hmod_cent$b),
         height_pred = a + b * (weight - xbar)) %>% 
  ggplot(aes(weight, height)) + 
  geom_point() +
  geom_line(aes(weight, height_pred), color = "green4") + 
  labs(title = "Centered Model")

```

Fit it without the centering

```{r}
hmod <- 
  quap(
    alist(
      height ~ dnorm(mu, sigma),
      mu <- a + b*(weight),
      a ~ dnorm(178, 20),
      b ~ dlnorm(0, 1),
      sigma ~ dunif(0, 50)
    ),
    data = d2)

post_hmod <- extract.samples(hmod) #--get our posterior samples of a, b, and sigma

p_hmod <- 
  d2 %>% 
  mutate(a = mean(post_hmod$a),
         b = mean(post_hmod$b),
         height_pred = a + b * (weight)) %>% 
  ggplot(aes(weight, height)) + 
  geom_point() +
  geom_line(aes(weight, height_pred), color = "red") + 
  labs(title = "Un-centered Model")

```

```{r}
library(patchwork)
p_hmod_cent + p_hmod
```

It looks identical to me. Look at the covariance, they say.

The 'centered' model:
```{r}
precis(hmod_cent)
```

The un-centered one:
```{r}
precis(hmod)
```

Centered:
```{r}
cov2cor(vcov(hmod_cent))
```

The 'un-centered' model:
```{r}
cov2cor(vcov(hmod))
```

I would like help.


### M8
More knots to the cherry blossom data!

Start with book example
```{r}
data("cherry_blossoms")
cd <- cherry_blossoms %>% filter(!is.na(doy))
num_knots <- 15
knot_list <- quantile( cd$year, probs = seq(0, 1, length.out = num_knots))
```

Fit the splines
```{r}
library(splines)
splB15 <- bs(cd$year,
            knots = knot_list[-c(1, num_knots)],
            degree = 3, intercept = TRUE)

plot (NULL, xlim = range(cd$year), ylim = c(0, 1), xlab = "year", ylab = "basis")
for (i in 1:ncol(splB15)) lines(cd$year, splB15[, i])

```

Well I would use ggplot2 anyways. Let's try using the geom_smooth functionality.
**Note** The default degree for *bs* is 3. If I try specifying *knots* as *num_knots*, it does something strange. So instead I specify *df* as *num_knots*, and it 'works'. 

```{r}
ggplot(cd, aes(year, doy)) +
  geom_point() +
  geom_smooth(method = lm, formula = y ~ splines::bs(x, df = num_knots, degree = 3)) + 
  labs(title = "15 knots")
```

```{r}
ggplot(cd, aes(year, doy)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = lm, formula = y ~ splines::bs(x, df = num_knots*2, degree = 3), color = "red") + 
  labs(title = "30 knots")
```

```{r}
ggplot(cd, aes(year, doy)) +
  geom_point() +
  geom_smooth(method = lm, formula = y ~ splines::bs(x, num_knots))
```

Use quap, not sure how to translate this plot into ggplot2. It seems the *se* argument in *geom_smooth* matches these though.  
```{r}
spl_mod <- 
  quap(
    alist(
      D ~ dnorm(mu, sigma),
      mu <- a + B %*% w, 
      a ~ dnorm(100, 10),
      w ~ dnorm(0, 10),
      sigma ~ dexp(1)
    ), data = list(D = cd$doy, B = splB15),
    start = list(w = rep(0, ncol(splB15)))
  )

mu <- link(spl_mod)
mu_PI <- apply(mu, 2, PI, 0.97)
plot(cd$year, cd$doy)
shade(mu_PI, cd$year)

```

