---
title: "Those Poor Manatees"
author: "Gina Nichols"
date: "7/28/2020"
output: html_document
---


```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(cowplot)
library(emo)
library(rethinking)
theme_set(theme_bw())
```

# Easy

## 8E1. 
**Name a hypothetical 3rd variable that could lead to an interaction**
- bread dough rising *FERNANDO THIS IS FOR YOU!* because of yeast could depend on the **temperature**
- education leading to higher income, could depend on **your socio-economic status before your eduction**
- gasoline makes a car go, could depend on *the status of the gas pedal*


## 8E2. 
**Which explanation invokes an interaction?**
4 is the only one that doesn't, I think

## 8E3
$onion\ carmelization = \alpha + \beta_{heat}(Heat) + \beta_{mois}(Mois) + \beta_{heat,moist}(Heat)(Mois)$
That's enough of that.

# Medium

## 8M1
**Add temperature to the tulips (shade, water)**
The main effects of shade and water, as well as their interaction, depend on the temperature. 

## 8M2
**What is an example of a masked relationship?**
I'm struggling with these. 

## 8M3
**Invent a reression that makes bloom size 0 if the temperature is hot**
$relative\_bloom\_size_i = (1 - Heatstress_i)(\alpha + \beta_{shade}(shade_i) + \beta_{water}(water_i) + \beta_{shade,water}(shade_i)(water_i)$

## 8M4
**Force water effect to be positive and shade effect to be negative.**
Remember rlnorm can be used to force things to be positive (page 96). 


```{r}
data(tulips)  
dt <- tulips %>% 
  as_tibble() %>% 
  mutate(blooms_std = blooms / max(blooms), #-don't standardize if 0 is meaninful. Like with ruggedness, only scale it
         water_cent = water - mean(water),
         shade_cent = shade - mean(shade))

dt

```


Try original priors first:
```{r normal priors tuplips}

m_tul1 <- quap(
  alist(
    blooms_std <- dnorm(mu, sigma),
    mu <- a + bw*water_cent + bs*shade_cent + bws*water_cent*shade_cent,
    a ~ dnorm(0.5, 0.25),
    bw ~ dnorm(0, 0.25),
    bs ~ dnorm(0, 0.25),
    bws ~ dnorm(0, 0.25),
    sigma ~ dexp(1)
  ), data = dt)

```


```{r tulip prior function, echo=F}

f_shade_priors <- function(mymodel = m_tul1, mydata = dt, mylines = 50, mycolor = "pink"){

prior <- extract.prior(mymodel)
    
  #shade = -1 ##################################################
  #--for the actual points on the graph
ds1 <- dt %>%
  select(water_cent, shade_cent, blooms_std) %>%
  filter(shade_cent == -1)

#--keep only 50 lines
mu_s1 <- link(mymodel, post = prior, data = data.frame(shade_cent = -1, water_cent = -1:1))[1:mylines,]

p1 <- t(mu_s1) %>% 
  as_tibble() %>% 
  mutate(water_cent = c(-1, 0, 1)) %>% 
  pivot_longer(-water_cent) %>% 
  ggplot() +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0, linetype = "dashed") + 
  geom_point(data = ds1,
             aes(x = water_cent, y = blooms_std)) +
  geom_line(aes(x = water_cent, y = value, group = name), color = mycolor) +
  labs(title = "shade = -1")


  #shade = 0 ##################################################
  #--for the actual points on the graph
ds2 <- dt %>%
  select(water_cent, shade_cent, blooms_std) %>%
  filter(shade_cent == 0)

#--keep only 50 lines
mu_s2 <- link(mymodel, post = prior, data = data.frame(shade_cent = 0, water_cent = -1:1))[1:mylines,]

p2 <- t(mu_s2) %>% 
  as_tibble() %>% 
  mutate(water_cent = c(-1, 0, 1)) %>% 
  pivot_longer(-water_cent) %>% 
  ggplot() +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0, linetype = "dashed") + 
  geom_point(data = ds2,
             aes(x = water_cent, y = blooms_std)) +
  geom_line(aes(x = water_cent, y = value, group = name), color = mycolor) +
  labs(title = "shade = 0")


  #shade = 1 ##################################################
  #--for the actual points on the graph
ds3 <- dt %>%
  select(water_cent, shade_cent, blooms_std) %>%
  filter(shade_cent == 1)

#--keep only 50 lines
mu_s3 <- link(mymodel, post = prior, data = data.frame(shade_cent = 1, water_cent = -1:1))[1:mylines,]

p3 <- t(mu_s3) %>% 
  as_tibble() %>% 
  mutate(water_cent = c(-1, 0, 1)) %>% 
  pivot_longer(-water_cent) %>% 
  ggplot() +
   geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0, linetype = "dashed") + 
  geom_line(aes(x = water_cent, y = value, group = name), color = mycolor) +
   geom_point(data = ds3,
             aes(x = water_cent, y = blooms_std)) +
  labs(title = "shade = 1")

cowplot::plot_grid(p1, p2, p3, ncol = 3)

}

f_shade_priors()
```

Try adjusting priors. 
First just make water thing positive:

```{r tulip water pos}

m_tul2 <- quap(
  alist(
    blooms_std <- dnorm(mu, sigma),
    mu <- a + bw*water_cent + bs*shade_cent + bws*water_cent*shade_cent,
    a ~ dnorm(0.5, 0.25),
    bw ~ dlnorm(0, 0.25), #--make this always positive
    bs ~ dnorm(0, 0.25),
    bws ~ dnorm(0, 0.25),
    sigma ~ dexp(1)
  ), data = dt)

f_shade_priors(mymodel = m_tul2, mycolor = "purple")
```

Oof those slopes are way to dramatic. Try restricting them:

```{r tulip water pos2}

m_tul2a <- quap(
  alist(
    blooms_std <- dnorm(mu, sigma),
    mu <- a + bw*water_cent + bs*shade_cent + bws*water_cent*shade_cent,
    a ~ dnorm(0.5, 0.25),
    bw ~ dlnorm(0, 0.001), #--make this always positive, smaller range?
    bs ~ dnorm(0, 0.25),
    bws ~ dnorm(0, 0.25),
    sigma ~ dexp(1)
  ), data = dt)

f_shade_priors(mymodel = m_tul2a, mycolor = "purple")
```

Maybe it has to be in tandem with the shade being negative. 


First water positive and shade negative try #1:

```{r tulip wat pos shade neg}

m_tul3 <- quap(
  alist(
    blooms_std <- dnorm(mu, sigma),
    mu <- a + bw*water_cent - bs*shade_cent + bws*water_cent*shade_cent,
    a ~ dnorm(0.5, 0.25),
    bw ~ dlnorm(0, 0.25), #--make this always positive
    bs ~ dlnorm(0, 0.25), #--make this always pos, but it's subtracted (?)
    bws ~ dnorm(0, 0.25),
    sigma ~ dexp(1)
  ), data = dt)

f_shade_priors(mymodel = m_tul3, mycolor = "dodgerblue")
```

# HELP `r emo::ji("sweat")`