---
title: "Chapter 8 Stat Rethink"
author: "Anabelle Laurent"
date: "7/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rethinking)
data(tulips)
d <- tulips
str(d)

d$blooms_std <- d$blooms / max(d$blooms)
d$water_cent <- d$water - mean(d$water)
d$shade_cent <- d$shade - mean(d$shade)
```

### E.1
1) temperature will interact with yeast  
2) professional experience (in year) will interact with the income  
3) the car brand will interact with gasoline  

### E.2
1) dryness will depend of the heat  
2) regarding the statement, it does not seems there is an interaction between cylinders and injector  
4) interaction between appendage and social skills to predict animal species intelligence

### E.3
```{r,eval=FALSE}
 mu <- a + bh*heat_i + bd*dryness_i + bhd*heat_i*dryness_i # for the example 1
 mu <- a + bs*social_i + ba*social_i + bhd*heat_i*appendage_i # for the example 4
```

### M.1
Bloom depends on temperature, water and shade. 3-way interaction. We learnt that bloom depend on water and shade at low temperature. There is an effect of high temperature too

### M.2
```{r eval=FALSE}
 blooms_std ~ dnorm( mu , sigma )
        mu <- a + bw*water_cent + bs*shade_cent + bt*temp_cent+ 
          bwst*water_cent*shade_cent*temp_cent +
          bws*water_cent*shade_cent +
          bwt*water_cent*temp_cent +
          bst*shade_cent*temp_cent
```

### M.3
```{r eval=FALSE}
raven_populationsize ~ dnorm( mu , sigma )
        mu <- a + bw*wolf_pop + bf*wolf_prey +  bwf*wolf_pop*wolf_prey
```

### M.4 
```{r}
m8.5 <- quap(
    alist(
        blooms_std ~ dnorm( mu , sigma ) ,
        mu <- a + bw*water_cent + bs*shade_cent + bws*water_cent*shade_cent ,
        a ~ dnorm( 0.5 , 0.25 ) ,
        bw ~ dnorm( 0 , 0.25 ) ,
        bs ~ dnorm( 0 , 0.25 ) ,
        bws ~ dnorm( 0 , 0.25 ) ,
        sigma ~ dexp( 1 )
    ) , data=d )
precis(m8.5)

set.seed(7)
prior <- extract.prior(m8.5)

par(mfrow=c(1,3)) # 3 plots in 1 row
for ( s in -1:1 ) {
    idx <- which( d$shade_cent==s )
    plot( d$water_cent[idx] , d$blooms_std[idx] , xlim=c(-1,1) , ylim=c(0,1) ,
        xlab="water" , ylab="blooms" , pch=16 , col=rangi2 )
    mu <- link( m8.5 ,post=prior, data=data.frame( shade_cent=s , water_cent=-1:1 ) )
    for ( i in 1:20 ) lines( -1:1 , mu[i,] , col=col.alpha("black",0.3) )
}
```
Question: should we use a uniform prior or half-student distribution to constrain the prior to be positive/negative?  

### H.1 
```{r}
m8.6 <- quap(
    alist(
        blooms_std ~ dnorm( mu , sigma ) ,
        mu <- a + bbed[bed] + bw*water_cent + bs*shade_cent + bws*water_cent*shade_cent ,
        a ~ dnorm( 130 , 100 ) , # 130 = mean of bloom in the datasets 
        bbed[bed] ~ dnorm( 0 , 0.25 ) ,  
        bw ~ dnorm( 0 , 0.25 ) ,
        bs ~ dnorm( 0 , 0.25 ) ,
        bws ~ dnorm( 0 , 0.25 ) ,
        sigma ~ dexp( 1 )
    ) , data=d )
precis(m8.6,depth=2)
```
### H2
```{r}
set.seed(77)
compare( m8.5, m8.6 , func=WAIC )
```

### H3
```{r}
data(rugged)
d <- rugged
d$log_gdp <- log(d$rgdppc_2000)
dd <- d[ complete.cases(d$rgdppc_2000) , ]
dd$log_gdp_std <- dd$log_gdp / mean(dd$log_gdp)
dd$rugged_std <- dd$rugged / max(dd$rugged)
dd$cid <- ifelse( dd$cont_africa==1 , 1 , 2 )

m8.3 <- quap(
    alist(
        log_gdp_std ~ dnorm( mu , sigma ) ,
        mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
        a[cid] ~ dnorm( 1 , 0.1 ) ,
        b[cid] ~ dnorm( 0 , 0.3 ) ,
        sigma ~ dexp( 1 )
    ) , data=dd )
precis( m8.3 , depth=2 )
plot( PSIS( m8.3 , pointwise=TRUE )$k )
```
what if we remove Seychelles?
```{r}
d2 <- dd[dd$country != "Seychelles", ]
m8.3bis <- quap(
    alist(
        log_gdp_std ~ dnorm( mu , sigma ) ,
        mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
        a[cid] ~ dnorm( 1 , 0.1 ) ,
        b[cid] ~ dnorm( 0 , 0.3 ) ,
        sigma ~ dexp( 1 )
    ) , data=d2 )
precis( m8.3bis , depth=2 )
compare( m8.3,m8.3bis , func=WAIC )
```

```{r}
m8.3stud <- quap(
    alist(
        log_gdp_std ~ dstudent(2, mu , sigma ) ,
        mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
        a[cid] ~ dnorm( 1 , 0.1 ) ,
        b[cid] ~ dnorm( 0 , 0.3 ) ,
        sigma ~ dexp( 1 )
    ) , data=dd )
precis( m8.3stud , depth=2 )
compare( m8.3,m8.3stud , func=WAIC )
```

