---
title: "Stat_rethink Chapter 8"
author: "Mariana CHiozza"
date: "7/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Conditional Manatees


## Building an interaction

AIC criterium: 

Only for comparing a model that treats all continents the same way vs a model that allows different slopes in different continents. We need models that use all of the same data (Chaper 7).


If we want to include continents in the model: Instead of doing separate linear regressions for each continent (African vs Non-African), include them as having different intercepts (alphas) using an indexing approach. We can use the same priors or assign different priors to the different continents.

NOTE: Adding a term to the linear model with 0/1 values for African/Non-African will not work because it will tell the model that mu for Africa nations is more uncertain than mu for Non-African nations.

BUT, this is going to show us the same slope. Including continents as another intercept is allowing the model to predict a different (in this case lower) mean for African (0.88) than for Non-African (1.05) nations (Figure 8.4, pag. 247 and output from 8.10 code).


How to make the slope conditional on continent to include the changes in slope due to different continents: Use the indexing approach as with the intercepts (and use the same priors).

This way, the continents have 0.13 slope for African and -0.14 for Non-African nations.


## Continuous interaction

```{r}
library(rethinking)
data(tulips)
d<-tulips
```

```{r}
d$blooms_std<-d$blooms/max(d$blooms)
d$water_cent<-d$water - mean(d$water)
d$shade_cent<-d$shade - mean(d$shade)

m8.5<-quap(
  alist(
    blooms_std~dnorm(mu, sigma),
    mu<- a + bw*water_cent + bs*shade_cent + bws*water_cent*shade_cent,
    a ~ dnorm(0.5 , 0.25),
    bw ~ dnorm (0, 0.25) ,
    bs ~ dnorm(0, 0.25),
    bws~ dnorm(0, 0.25),
    sigma ~ dexp (1)
  ), data = d
)

```






### Practice

## 8E1.

1) Ambient temperature: Optimum-supra-optimum and sub-optimum temperatures 

2) Countries: Developed vs Developing countries

3) Type of car: Gasoline only, hybrids, electric only.


## 8E2.

(3)

## 8E3.

(1)

y= onion caramelization
x= cooking heat

y= m*x + b

(2)

y= car speed
x= number of cilinders or injectors

y= m*x + b

(3)

y= republican people
x= republican parents/friends

y= m[cid]*x + b[cid]

(4)

y= animal inteligence
x= social skils or manipulative appendages

y= m*x + b


## 8M1.

Plant development is affected by temperature. There is an optimun air temperature at which bloom will occur. Temperatures below or above that optimun will not produce any flower. 

Also, high temperatures could also increase soil water evaporation and diminish the available water for growing. And, shading could decrease the canopy temperature below the optimun.

## 8M2.

T = Temperature (1= low, 0=high, 2=optimun)

B~Normal (mu, sigma)

mu= alpha + bT*T = 0

## 8M3.

Ravens population may rise as the number of wolves increase due to the availability of food, up to a point where the population of ravens will not longer depend on the availability of food.






