---
title: "Stat Rethink Chap 13"
author: "Anabelle Laurent"
date: "9/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE)
library(rethinking)
library(tidyverse)
library(cowplot)
library(emo)
theme_set(theme_bw())
```


### Note 

```{r, results="hide"}
data(reedfrogs)
d <- reedfrogs

# make the tank cluster variable
d$tank <- 1:nrow(d)

dat <- list(
    S = d$surv,
    N = d$density,
    tank = d$tank )

# approximate posterior
m13.1 <- ulam(
    alist(
        S ~ dbinom( N , p ) ,
        logit(p) <- a[tank] ,
        a[tank] ~ dnorm( 0 , 1.5 )
    ), data=dat , chains=4 , log_lik=TRUE )

## R code 13.3
m13.2 <- ulam(
    alist(
        S ~ dbinom( N , p ) ,
        logit(p) <- a[tank] ,
        a[tank] ~ dnorm( a_bar , sigma ) ,
        a_bar ~ dnorm( 0 , 1.5 ) ,
        sigma ~ dexp( 1 )
    ), data=dat , chains=4 , log_lik=TRUE )
```
```{r}
compare( m13.1 , m13.2 )
```

Remember the number of effective number is like the penalty term (called p_waic) (see p.220).  
Regularizing prior to help the model not get too exited about the data.  
**Shrinkage** results from regularization and **pooling** each cluster provides information that can be use to improve the estimate of other clusters.  
sigma represents the variation across clusters.  

Varying intercepts (from using a multilevel model) are just regularized estimates by estimating how diverse the clusters are while estimating the features of each cluster. 


### Practice

## E1 


```{r}
library(ggtext)
ggplot(data = data.frame(x = c(-3, 3)), aes(x)) +
  stat_function(fun = dnorm, n = 101, args = list(mean = 0, sd = 1)) + ylab("") +
  scale_y_continuous(breaks = NULL) +
  stat_function(fun = dnorm, n = 101, args = list(mean = 0, sd = 2),col="red") + ylab("") +
  labs(
    title = "sd=1 in black ; sd=2 in red"
  )
```
The priors with the smaller sd will produce more shrinkage (sd=1)  

## E2 & E3
```{r,eval=FALSE}
m1<- ulam(
    alist(
        y ~ dbinom( 1 , p ) ,
        logit(p) <- a[group] + beta*X ,
        a[group] ~ dnorm( a_bar , sigma ) ,
        a_bar ~ dnorm( 0 , 1.5 ) ,
        sigma ~ dexp( 1 ),
        beta ~ dnorm( 0 , 0.05 )
    ), data=dat , chains=4 , log_lik=TRUE )


m2<- ulam(
alist(y~dnorm(mu,sigma),
      mu=a[group]+ b*x,
      a[group] ~ dnorm(a_bar, sigma_a),
      b ~ dnorm(0,1),
      a_bar ~ dnorm(0,1),
      sigma_a ~ dnorm(0,1)
), data=dat , chains=4 , log_lik=TRUE )
```


## E4 `r emo::ji("fish")`

```{r  eval=FALSE}
m3<- ulam(
  alist(
    T ~ dpois( lambda ),
    log(lambda) <- a[group],
    a_bar ~ dnorm(0,1),
    sigma_a ~ dnorm(0,1)
  ), data=dat , chains=4 , log_lik=TRUE )

```


## E5
```{r questionE5, eval=FALSE}
m4<-alist(
    T ~ dpois( lambda ),
    log(lambda) <- a[group] + g[block_id],
  g[block_id] ~ dnorm( 0 , sigma_g ),
        a ~ dnorm( 0 , 1.5 ),
        sigma_a ~ dexp(1),
        sigma_g ~ dexp(1)

```

## M1
```{r questionM1, results="hide"}
d <- reedfrogs
d$tank <- 1:nrow(d)
d$predation <- ifelse(test = d$pred == "pred", yes = 1, no = 0)
d$frogsize <- ifelse(test = d$size == "big", yes = 1, no = 0)

dat <- list(
    S = d$surv,
    N = d$density,
    tank = d$tank,
    P=d$predation,
    FS=d$frogsize)


M1 <- ulam(
  alist(
    S ~ dbinom( N , p ) ,
    logit(p) <- a[tank] + bP*P ,
    a[tank] ~ dnorm( a_bar , sigma ) ,
    a_bar ~ dnorm( 0 , 1.5 ) ,
    sigma ~ dexp( 1 ),
    bP ~ dnorm(0,5)
  ), data=dat , iter=4000 , chains=4 , log_lik=TRUE )

M1_2 <- ulam(
  alist(
    S ~ dbinom( N , p ) ,
    logit(p) <- a[tank] + bS*FS ,
    a[tank] ~ dnorm( a_bar , sigma ) ,
    a_bar ~ dnorm( 0 , 1.5 ) ,
    sigma ~ dexp( 1 ),
    bS ~ dnorm(0,5)
  ), data=dat , iter=4000 , chains=4 , log_lik=TRUE )


M1_both <- ulam(
  alist(
    S ~ dbinom( N , p ) ,
    logit(p) <- a[tank] + bS*FS + bP*P ,
    a[tank] ~ dnorm( a_bar , sigma ) ,
    a_bar ~ dnorm( 0 , 1.5 ) ,
    sigma ~ dexp( 1 ),
    bS ~ dnorm(0,5),
    bP ~ dnorm(0,5)
  ), data=dat , iter=4000 , chains=4, log_lik=TRUE  )

M1_int <- ulam(
  alist(
    S ~ dbinom( N , p ) ,
    logit(p) <- a[tank] + bS*FS + bP*P + bSP*P*FS,
    a[tank] ~ dnorm( a_bar , sigma ) ,
    a_bar ~ dnorm( 0 , 1.5 ) ,
    sigma ~ dexp( 1 ),
    bS ~ dnorm(0,5),
    bP ~ dnorm(0,5),
    bSP ~ dnorm(0,5)
  ), data=dat , iter=4000 , chains=4, log_lik=TRUE  )
```

```{r}
compare(M1,M1_2,M1_both,M1_int)
```


## M3

```{r results="hide"}
M3 <- ulam(
    alist(
        S ~ dbinom( N , p ) ,
        logit(p) <- a[tank] ,
        a[tank] ~ dcauchy( a_bar , sigma ) ,
        a_bar ~ dnorm( 0 , 1 ) ,
        sigma ~ dexp( 1 )
    ), data=dat , chains=4 , log_lik=TRUE )

# model from the book
m13.2 <- ulam(
    alist(
        S ~ dbinom( N , p ) ,
        logit(p) <- a[tank] ,
        a[tank] ~ dnorm( a_bar , sigma ) ,
        a_bar ~ dnorm( 0 , 1.5 ) ,
        sigma ~ dexp( 1 )
    ), data=dat , chains=4 , log_lik=TRUE )

```

```{r}
precis(m13.2,depth=2)
plot(precis(m13.2,depth=2))

precis(M3,depth=2)
plot(precis(M3,depth=2))

compare( M3 , m13.2 )
```

```{r}
post_normal <- extract.samples(m13.2)
a_normal <- apply(X = post_normal$a, MARGIN = 2, FUN = mean)
post_cauchy <- extract.samples(M3)
a_cauchy <- apply(X = post_cauchy$a, MARGIN = 2, FUN = mean)
plot( a_normal , a_cauchy , pch=16 , col=rangi2 ,
      xlab="Gaussian prior" , ylab="Cauchy prior" )
```

Shrinkage using cauchy  

## M4
```{r results="hide"}
M4 <- ulam(
    alist(
        S ~ dbinom( N , p ) ,
        logit(p) <- a[tank] ,
        a[tank] ~ dstudent(2, a_bar , sigma ) ,
        a_bar ~ dnorm( 0 , 1 ) ,
        sigma ~ dexp( 1 )
    ), data=dat , chains=4 , log_lik=TRUE )
plot(precis(M4,depth=2))
```


## H1 `r emo::ji("pregnant")`

```{r}
data(bangladesh)
d <- bangladesh
str(d)

d$district_id <- as.integer(as.factor(d$district))
sort(unique(d$district_id))

dat <- list(
    C = d$use.contraception,
    dist =  d$district_id)
```

```{r results="hide"}

H1 <- ulam(
  alist(
    C ~ dbinom(1, p) ,
    logit(p) <- a[dist],
    a[dist] ~ dnorm(a_bar, sigma),
    a_bar ~ dnorm(0, 5),
    sigma ~ dexp(1)
  ), data=dat, chains=4 , log_lik=TRUE)


H1_nomulti <- ulam(
  alist(
    C ~ dbinom(1, p) ,
    logit(p) <- a[dist],
    a[dist] ~ dnorm(0, 1.5)
  ), data=dat, chains=4 , log_lik=TRUE)

```


```{r}
compare(H1,H1_nomulti)
```

```{r}
# extract Stan samples
post <- extract.samples(H1)
dat$propcontra.est <- logistic( apply( post$a , 2 , mean ) )
post2 <- extract.samples(H1_nomulti)
dat$propcontra.est2 <- logistic( apply( post2$a , 2 , mean ) )

rawdata<- d %>% group_by(district_id) %>% summarise(raw_c=sum(use.contraception/n()))

# display raw contraception rate
plot( rawdata$raw_c , ylim=c(0,1) , pch=16 , xaxt="n" ,
    xlab="district" , ylab="proportion using contraception" , col="black" )
axis( 1 , at=c(1,16,32,48) , labels=c(1,16,32,48) )

# posterior means
points( dat$propcontra.est,col="blue",pch=16) # red for multimodel
points( dat$propcontra.est2,col="red",pch=16) # red for model fixed effect 
legend("topright",legend=c("data","fixed effects", "multilevel"),fill=c("black","red","blue"))

# mark posterior mean probability across districts
abline( h=mean(inv_logit(post$a_bar)) , lty=2 )


```

