---
title: "Stat_rethink Chapter 13"
author: "Mariana CHiozza"
date: "9/01/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Models with Memory

Cafes example: The distribution of waiting times in the population is the prior for each cafe. Here we are not using the posterior distribution of one given cafe as a prior for the next cafe because this implies that the two cafes have the same average waiting time.


## Practice Problems

### 13E1.

The prior with the bigger sigma will produce more shrinkage in the estimates since it has a wider distribution.


### 13E2.

m13E2 <- ulam(
    alist(
        y ~ dbinom( 1 , p ) ,
        logit(p) <- a[group] + b*x ,
        b ~ dnorm( 0 , 0.5 ),
    ## adaptive priors
        a[group] ~ dnorm( a_bar , sigma_a ),
    ## hyper-priors
        a_bar ~ dnorm( 0 , 1.5 ),
        sigma_a ~ dexp(1)
    


### 13E3.


m13E3 <- ulam(
    alist(
        y ~ dnorm( mu , sigma) ,
        mu<-a[group] + b*x ,
        b ~ dnorm( 0 , 1 ),
        a[group] ~ dnorm( a_bar , sigma_a ),
        a_bar ~ dnorm( 0 , 5 ),
        sigma_a ~ dexp(1),
        sigma   ~ dexp(1)


### 13E4.


### 13E5.


### 13M1.





```{r}
library(rethinking)
data(reedfrogs)
d <- reedfrogs
str(d)

## R code 13.2
# make the tank cluster variable
d$tank <- 1:nrow(d)

dat <- list(
    S = d$surv,
    N = d$density,
    P = d$pred,
    Z = d$size,
    tank = d$tank )

# approximate posterior

m13M1.1 <- ulam(
    alist(
        S ~ dbinom( N , p ) ,
        logit(p) <- a[tank] ,
        a[tank] ~ dnorm( a_bar , sigma ) ,
        a_bar ~ dnorm( 0 , 1.5 ) ,
        sigma ~ dexp( 1 )
    ), data=dat , chains=4 , log_lik=TRUE )



# Should we add pred variable with a binomial distribution????
# m13M1.2 <- ulam(
#    alist(
  #      S ~ dbinom( N , p ) ,
  #     logit(p) <- a[tank] + b[pred],
  #      b[pred] ~ dbinom(??),
        
        
  #      a[tank] ~ dnorm( a_bar , sigma ) ,
  #      a_bar ~ dnorm( 0 , 1.5 ) ,
  #      sigma ~ dexp( 1 )
  #      ), data=dat , chains=4 , log_lik=TRUE )


```

