---
title: "chapter4"
author: "Anabelle Laurent"
date: "6/16/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### E.1 
first line is the likelihood as it is the distribution assigned ton an observed variable (see p.33)  

### E.2 
two parameters, mu and sigma (unobservable things)    

### E.3  
I used the "overthinking" note page 78  
Pr(mu,sigma|Yij) = N/D  
N= Normal(Yij|mu,sigma) x Normal(0,10) x Uniform(0,10)  
D= integral(Normal(Yij|mu,sigma)) x Normal(0,10) x Uniform(0,10)  

### E.4  
second line = linear model  

### E.5  
Three parameters alpha, beta and sigma

### M.1
```{r}
mu<-rnorm(100,0,10)
sigma<-runif(100,0,10)
y<-rnorm(100,mu,sigma)
print(y)
```

### M.4
$$
y_{i} \sim Normal(\mu_i, \sigma) 
$$

$$
\mu_i = \alpha + \beta x_i
$$

$$
\alpha \sim Normal(130,10)   
$$
$$
\beta \sim Normal(3,2)
$$

$$
\sigma \sim Uniform(0,40)
$$

i = row index in a data frame  
first line = likelihood  
second line = linear model  
then it's the priors  
They are students, I don't know the age but I'am assuming they are about 10 years old at the beginning of the study (affects the prior alpha)  
alpha is the intercept and beta is the slope (kind of rate)  
sigma is the standard deviation, it must be positive. I choose a flat prior. 

### M.5
As student got taller each year we need to inform the prior for the slope (beta)  

### M.6
It impacts sigma