---
title: "Chapter 5 Stat Rethink Practice"
author: "Anabelle Laurent"
date: "6/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rethinking)
```

## E1
Models (2), (3) and (4) have two predictors $x_i$ and $z_i$  

## E2
$$
D_{i} \sim Normal(\mu_i, \sigma) 
$$
$$
\mu_i = \alpha + \beta_L* L_i + \beta_P* P_i
$$
with:  
D is the animal diversity  
L is the latitude (predictor)  
P is the plant diversity (predictor)  
$\alpha$ is the mean  
$\beta_L$ and $\beta_P$ are slope parameters  

## E3
$$
T_{i} \sim Normal(\mu_i, \sigma) 
$$
$$
\mu_i = \alpha + \beta_F* F_i + \beta_S* S_i
$$
with:  
T is the time to PhD degree
F is the amount of funding (predictor)  
S is the size of laboratory (predictor)  
$\alpha$ is the mean  
$\beta_F$ and $\beta_S$ are slope parameters  

Each slope parameters ($\beta_F$ and $\beta_S$) are on the right side of zero (positive slope)  

## E4
models (3) and (4)

## M1 Spurious correlation
It's like the model m5.3 (page 133)

```{r}
data(WaffleDivorce)
d <- WaffleDivorce

# standardize variables
d$D <- standardize( d$Divorce )
d$M <- standardize( d$Marriage )
d$A <- standardize( d$MedianAgeMarriage )

m1 <- quap(
    alist(
        D ~ dnorm( mu , sigma ) ,
        mu <- a + bM*M + bA*A ,
        a ~ dnorm( 0 , 0.2 ) ,
        bM ~ dnorm( 0 , 0.5 ) ,
        bA ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data = d )
precis( m1 )
plot( coeftab(m1), par=c("bA","bM") )
```

## M2 Masked relationship
It's like the model m5.6 (page 148)

```{r}
data(milk)
d <- milk
d$K <- standardize( d$kcal.per.g )
d$N <- standardize( d$neocortex.perc )
d$M <- standardize( log(d$mass) )
dcc <- d[ complete.cases(d$K,d$N,d$M) , ]
m5.6 <- quap(
    alist(
        K ~ dnorm( mu , sigma ) ,
        mu <- a + bM*M ,
        a ~ dnorm( 0 , 0.2 ) ,
        bM ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data=dcc )
precis(m5.6)
```
## M3
How a higher divorce rate cause a higger marriage rate? Because a higger divorce rate can mean a higher number of inhabitants (so a higher rate of marriage)

## M4
```{r, eval=FALSE}
m1 <- quap(
    alist(
        D ~ dnorm( mu , sigma ) ,
        mu <- a + bM*M + bA*A + bLDS*LDS,
        a ~ dnorm( 0 , 0.2 ) ,
        bM ~ dnorm( 0 , 0.5 ) ,
        bA ~ dnorm( 0 , 0.5 ) ,
        bLDS ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data = d )
```

## M5
$$
O_{i} \sim Normal(\mu_i, \sigma) 
$$
$$
\mu_i = \alpha + \beta_G* G_i + \beta_D* D_i + \beta_E* E_i
$$

With:  
O = obesity rate  
G = price of gasoline  
D = time driving  
E = time exercising  

Question: How should we measure the variables D and E? Time? Frequence per week? Energy (calories)?  

## H1

## H2
```{r}
data(WaffleDivorce)
d <- list()
d$A <- standardize( WaffleDivorce$MedianAgeMarriage )
d$D <- standardize( WaffleDivorce$Divorce )
d$M <- standardize( WaffleDivorce$Marriage )

m5.3_A <- quap(
    alist(
      ## A -> M -> D
        D ~ dnorm( mu , sigma ) ,
        mu <- a + bM*M + bA*A ,
        a ~ dnorm( 0 , 0.2 ) ,
        bM ~ dnorm( 0 , 0.5 ) ,
        bA ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 ),
      ## A -> M
        M ~ dnorm( mu_M , sigma_M ),
        mu_M <- aM + bAM*A,
        aM ~ dnorm( 0 , 0.2 ),
        bAM ~ dnorm( 0 , 0.5 ),
        sigma_M ~ dexp( 1 )
    ) , data = d )
```


## H4
```{r}
data(WaffleDivorce)
d <- WaffleDivorce
str(d$South)
# standardize variables
d$D <- standardize( d$Divorce )
d$M <- standardize( d$Marriage )
d$A <- standardize( d$MedianAgeMarriage )
d$S <- standardize( d$South )
m5 <- quap(
    alist(
        D ~ dnorm( mu , sigma ) ,
        mu <- a + bM*M + bA*A + c[South]*S,
        a ~ dnorm( 0 , 0.2 ) ,
        bM ~ dnorm( 0 , 0.5 ) ,
        bA ~ dnorm( 0 , 0.5 ) ,
        c[South] ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data = d )
precis(m5,depth = 2)
#plot(coeftab(m5),par=c("bM","bA","bS"))
```
```{r}
ggplot(data = WaffleDivorce, 
       aes(x = South, y =Divorce)) + geom_point()
```
```{r}
post <- extract.samples(m5)
post$diff_fm <- post$c[,1] - post$c[,2]
precis( post , depth=2 )
```

